---
title: HW3
author: Miracle Ephraim
format:
  pdf:
    documentclass: article
---

```{r setup, include=FALSE}


# install.packages("readr")
# install.packages("knitr")
# install.packages("rmarkdown")
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, data.table, gdata)


library(readr)

load("C:/Users/mirac/Documents/GitHub/econ470_ma/hw3/submission2/results/hw_workspace.Rdata")

```

# Question 1

Present a bar graph showing the proportion of states with a change in their cigarette tax in each year from 1970 to 1985.

```{r, echo=FALSE}

#| echo: false
ggplot(data=prop_tax, aes(Year,prop)) +
    geom_col() 

```


# Question 2

Plot on a single graph the average tax (in 2012 dollars) on cigarettes and the average price of a pack of cigarettes from 1970 to 2018.

```{r, echo=FALSE}

#| echo: false
ggplot() +
geom_line(data = avgs2, aes(Year, avg_tax), color = "blue") +
geom_line(data = avgs2, aes(Year, avg_cost), color = "light blue") 

```


# Question 3

Identify the 5 states with the highest increases in cigarette prices (in dollars) over the time period. Plot the average number of packs sold per capita for those states from 1970 to 2018.

```{r, echo=FALSE}

#| echo: false

top5 <- df %>%
    filter(Year == 1970 | Year == 2018) %>%
    group_by(state) %>%
    summarise(diff = cost_per_pack - lag(cost_per_pack)) %>%
    filter(!is.na(diff)) %>%
    arrange(desc(diff))

head(top5, 5)

ggplot(df %>% filter(state_abb == c("NY", "CT", "DC", "RI", "MA")), aes(Year, sales_per_capita, colour = state)) +
geom_line() 


```


# Question 4

Identify the 5 states with the lowest increases in cigarette prices over the time period. Plot the average number of packs sold per capita for those states from 1970 to 2018.

```{r, echo=FALSE}

#| echo: false

low5 <- df %>%
    filter(Year == 1970 | Year == 2018) %>%
    group_by(state) %>%
    summarise(diff = cost_per_pack - lag(cost_per_pack)) %>%
    filter(!is.na(diff)) %>%
    arrange(diff) 

head(low5, 5)

ggplot(df %>% filter(state_abb == c("MO", "ND", "TN", "GA", "NC")), aes(Year, sales_per_capita, colour = state)) +
geom_line() 

```


# Question 5

Compare the trends in sales from the 5 states with the highest price increases to those with the lowest price increases.

In states with higher prices, average sales decrease at a much faster rate compared to states where prices are lower.All states with high prices see sales per capita drop below 50 by the early 2000s, while states with lower reach that point by the end of follow-up (2018). These findings demonstrate the significant effect price has on cigarette sales, and are a good means of dissauding their use amongst the general population. 

# Question 6

Focusing only on the time period from 1970 to 1990, regress log sales on log prices to estimate the price elasticity of demand over that period. Interpret your results.

```{r, echo=FALSE}

summary(ols1)

```


On average, quantity demanded decreases by 17% for every 1% increase in price.


# Question 7

Again limiting to 1970 to 1990, regress log sales on log prices using the total (federal and state) cigarette tax (in dollars) as an instrument for log prices. Interpret your results and compare your estimates to those without an instrument. Are they different? If so, why?

```{r, echo=FALSE}

summary(step2)

```


With a coefficient of 0.5, we can say price of cigarettes is inelastic, with a 1% increase in price leading to only a 0.5% change in quantity demanded. These are quite different from the OLS estimates of elasticity, likely due to the endogeneity present in the OLS model. 


# Question 8

Show the first stage and reduced-form results from the instrument.

```{r, echo=FALSE}

summary(step1)

summary(step2)

```


# Question 9 

Repeat questions 1-3 focusing on the period from 1991 to 2015.

*OLS estimates*
```{r, echo=FALSE}

summary(ols2)

```

As price increases by 1%, quantity demanded decreases by 66%.


*IV estimates*
```{r, echo=FALSE}

summary(step4)

```

With a coefficient of -0.8, we can say price of cigarettes is inelastic, with a 1% increase in price leading to only a 0.8% change in quantity demanded. These are different from the OLS estimates of elasticity as well, but not as stark of a contrast from the OLS estimates.


*2SLS estimates*
```{r, echo=FALSE}

summary(step3)
summary(step4)

```


# Question 10 

Compare your elasticity estimates from 1970-1990 versus those from 1991-2015. Are they different? If so, why?

The estimates from each 20-year period are different from eachother, likely due to changes in attitudes around smoking and the increasing taxes placed on the product. 

**Github Repo:** [https://github.com/miracleephraim/hw3.git] (https://github.com/miracleephraim/hw3.git)